---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# od

<!-- badges: start -->
<!-- badges: end -->

The goal of od is to provide tools and example datasets for working with origin-destination (OD) datasets.

## Installation

<!-- You can install the released version of od from [CRAN](https://CRAN.R-project.org) with: 

``` r
install.packages("od")
```

And the development version from [GitHub](https://github.com/) with:

-->

``` r
# install.packages("devtools")
devtools::install_github("robinlovelace/od")
```

## Motivation

The package came out of conversations about the OD functions in the package `stplanr`, particularly `od2line` which works as follows:

```{r}
library(od) # load example datasets
od_data_leeds[1:2, 1:9] # OD data as data frame
od_data_zones[1:2, ]
desire_lines_stplanr = stplanr::od2line(od_data_leeds, od_data_zones)
desire_lines_stplanr[1:2, 1:9]
```

It works great, and is plenty fast enough for most applications, but there are some issues with `stplanr::od2line()`:

- The function is a commonly needed and low-level function, buried in a large package, reducing 'findability'
- To get the function you must install `stplanr` plus its numerous dependencies
- The function has not been optimised
- It has no class definition of 'od' data

The `od` package, as it currently stands, addresses the first three of these issues (it may at some point define a class for `od` objects but there are no immediate plans to do so).

The equivalent code in the `od` package is as follows:

```{r}
desire_lines_od = od_to_sfc(od_data_leeds, od_data_zones)
```

The result is an `sfc` object that has the same geometry as the output from `od2line`:

```{r}
desire_lines_od[1:2]
desire_lines_stplanr$geometry[1:2]
```

To make the results (almost) identical, we can specify `sf` outputs:

```{r}
desire_lines_od_sf1 = od_to_sf(od_data_leeds, od_data_zones)
desire_lines_od_sf1[1:2, 1:9]
sf::st_crs(desire_lines_od_sf1)
```

An additional option, that requires the `sf` package is to be installed, is to use sf during the creation of the lines, allowing crss to be provided:

```{r}
desire_lines_od_sf2 = od_to_sf(od_data_leeds, od_data_zones, package = "sf", crs = 4326)
sf::st_crs(desire_lines_od_sf2)
```


## Performance

The package is designed to be fast, with centroids only created when needed and the use of `sfheaders`.
The benchmark shows this performance.

```{r, message=FALSE}
bench::mark(check = FALSE, max_iterations = 100,
  stplanr = stplanr::od2line(od_data_leeds, od_data_zones),
  od = od_to_sfc(od_data_leeds, od_data_zones),
  od_sf1 = od_to_sf(od_data_leeds, od_data_zones),
  od_sf2 = od_to_sf(od_data_leeds, od_data_zones, package = "sf", crs = 4326)
)

```

```{r}
centroids = sf::st_centroid(od_data_zones)
bench::mark(check = FALSE, max_iterations = 100,
  stplanr_centroids = stplanr::od2line(od_data_leeds, centroids),
  od_sf3 = od_to_sf(od_data_leeds, centroids)
  # od_sf4 = od_to_sf(od_data_leeds, centroids, package = "sf", crs = 4326)
)

```


## Testing the similarity between stplanr and sf objects

Due to the way `od2line()` works and the outputs `sfheaders` functions, there are slight differences in the outputs of `od2line()` and `od_to_sf()`. 
These have no impact on real-world applications, as far as we are aware but are outlined below 

```{r}
desire_lines_od[1:2]
desire_lines_od_sf = sf::st_as_sf(od_data_leeds, geometry = desire_lines_od, crs = 4326)
identical(desire_lines_od_sf, desire_lines_stplanr)
identical(sf::st_drop_geometry(desire_lines_stplanr), sf::st_drop_geometry(desire_lines_od_sf))
attributes(desire_lines_od_sf$geometry) = attributes(desire_lines_stplanr$geometry)
identical(attributes(desire_lines_od_sf$geometry), attributes(desire_lines_stplanr$geometry))
attributes(desire_lines_od_sf) = attributes(desire_lines_stplanr)
identical(attributes(desire_lines_od_sf), attributes(desire_lines_stplanr))
identical(desire_lines_od_sf, desire_lines_stplanr)
identical(desire_lines_od_sf$geometry[1], desire_lines_stplanr$geometry[1])
identical(
  unclass(desire_lines_stplanr$geometry[1][[1]]),
  unclass(desire_lines_od_sf$geometry[1][[1]])
)
identical(
  unclass(desire_lines_stplanr$geometry[1][[1]]),
  unclass(desire_lines_od_sf$geometry[1][[1]])
)
attributes((desire_lines_od_sf$geometry[1]))
attributes((desire_lines_od_sf$geometry[1][[1]]))
attributes((desire_lines_stplanr$geometry[1][[1]]))
identical(
  as.numeric(unclass(desire_lines_stplanr$geometry[1][[1]])),
  as.numeric(unclass(desire_lines_od_sf$geometry[1][[1]]))
)

all.equal(
  as.numeric(unclass(desire_lines_stplanr$geometry[1][[1]])),
  as.numeric(unclass(desire_lines_od_sf$geometry[1][[1]]))
)
```

