---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# od

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/tic)](https://cran.r-project.org/package=tic)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-red.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Codecov test coverage](https://codecov.io/gh/itsleeds/od/branch/master/graph/badge.svg)](https://codecov.io/gh/itsleeds/od?branch=master)
[![R build status](https://github.com/itsleeds/od/workflows/R-CMD-check/badge.svg)](https://github.com/itsleeds/od/actions)
<!-- badges: end -->

The goal of od is to provide tools and example datasets for working with origin-destination (OD) datasets.

## Installation

<!-- You can install the released version of od from [CRAN](https://CRAN.R-project.org) with: 

``` r
install.packages("od")
```

And the development version from [GitHub](https://github.com/) with:

-->

``` r
# install.packages("devtools")
devtools::install_github("itsleeds/od")
```

## Motivation

The package originated as a set of functions in the package `stplanr` for working with origin-destination data.
The `od2line()` function, for example, takes a data frame and a spatial object as inputs and outputs geographic lines representing movement between origins and destinations:

```{r}
library(od) # load example datasets
od_data_df # OD data as data frame
od_data_centroids[1:2, ]
desire_lines_stplanr = stplanr::od2line(od_data_df, od_data_centroids)
desire_lines_stplanr[1:2, 1:9]
```

It works great, and is plenty fast enough for most applications, but there are some issues with `stplanr::od2line()` (which also affect the other `od_*()` functions in `stplanr`):

- The function is a commonly needed and low-level function, buried in a large package, reducing 'findability'
- To get the function you must install `stplanr` plus its numerous dependencies
- The function has not been optimised
- It has no class definition of 'od' data

The `od` package, as it currently stands, addresses the first three of these issues (it may at some point define a class for `od` objects but there are no immediate plans to do so).

The equivalent code in the `od` package is as follows:

```{r}
desire_lines_od = od_to_sfc(od_data_df, od_data_centroids)
```

The result is an `sfc` object that has the same geometry as the output from `od2line`:

```{r}
desire_lines_od[1:2]
desire_lines_stplanr$geometry[1:2]
```

These are 'desire lines' representing the shortest (straight line) path between two centoids and can plotted using geographic data and mapping packages such as `sf`, `mapview`, `tmap` and `mapdeck`, e.g.:

```{r, out.width="50%", fig.show='hold'}
plot(desire_lines_od)
plot(desire_lines_stplanr$geometry)
```

By default the package uses the `sfheaders` package to create sf objects.
You can can also specify `sf` outputs as follows:

```{r}
desire_lines_od_sf1 = od_to_sf(od_data_df, od_data_centroids)
```


## Performance

The package is designed to be fast, with centroids only created when needed and the use of `sfheaders`.

### Benchmark on a small dataset:

```{r, message=FALSE}
nrow(od_data_df)
bench::mark(check = FALSE, max_iterations = 100,
  stplanr = stplanr::od2line(od_data_df, od_data_zones),
  od = od_to_sfc(od_data_df, od_data_zones),
  od_sf1 = od_to_sf(od_data_df, od_data_zones),
  od_sf2 = od_to_sf(od_data_df, od_data_zones, package = "sf", crs = 4326)
)
```

## Related open source projects

- [stplanr](https://github.com/ropensci/stplanr/) is an R package package designed to support transport planning, with a focus on geographic transport datasets and many functions for working with OD data in the [od function family](https://docs.ropensci.org/stplanr/reference/index.html#section-work-with-od-data).
- [cartography](http://riatelab.github.io/cartography) is an R package with functions for working with OD data, including [`getLinkLayer()`](http://riatelab.github.io/cartography/docs/reference/getLinkLayer.html)
- [gravity](https://pacha.dev/gravity/) is an R package for developing 'gravity models' to estimate flow between zones
- [flowmap.gl](https://github.com/teralytics/flowmap.gl), a JavaScript package for visualising OD data
- [Arabesque](http://arabesque.ifsttar.fr/) is another JavaScript project for working with OD data

## Code of Conduct

Please note that the od project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
